Evolution of the AGN UV luminosity function from redshift 7.5

### 2017-09-21 Paper/paper.tex
Update text

\begin{keywords}
dark ages, reionization, firs stars -- intergalactic medium -- quasars: general -- galaxies: active 
\end{keywords}

%% * Possible outline of paper:
%% 
%% Introduction 
%% Data
%% Method
%%   Individual fits
%%   Global fits
%% Contribution to reionizatin 
%% Conclusions
%% 
%% * Optional possibilities:
%%
%% Bolometric LF
%% Public code description 
%% More global models: PLE, PDE, Trenti, White-Conroy
%% Goodness of fit 
%% BH growth 
%% Comparison with galaxy LF
%% Comparison with previous work

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.}
  \label{fig:e912}
\end{figure*}

@@

The faint-end slope $\beta$ seems to require a break in its evolution
at $z\sim 3$, as discussed in the previous section.  As a result, it
is difficult to model the evolution of $\beta$ using the Chebyshev
expansion of Equation~(\ref{eqn:cbs}).  Instead, we consider a double
power law model for $\beta$ and write
\begin{equation}
  f_3(1+z)=h+\frac{f_0}{10^{a*\zeta}+10^{b*\zeta}},
\end{equation}
where
\begin{equation}
  \zeta = \log_{10}\left(\frac{1+z}{1+z_0}\right),
\end{equation}
thus resulting in a five-parameter model with parameters $h$, $f_0$,
$z_0$, $a$ and $b$.  The parameters $a$ and $b$ thus determine the low
and high redshift slopes of this evolution, with a break at redshift
$z_0$.  This is similar to the model of \citet{2007ApJ...654..731H},
who also favoured a broken power law model for the evolution of the
faint-end slope of the bolometric luminosity function of quasars.

@@

@@

We assume that all quasars have a universal SED, which can be
parameterised in some fashion.  \citet{2015MNRAS.449.4204L} have
parameterised the quasar SED as a power law with a break at 912~{\AA},
\begin{equation}
f_\nu\propto\begin{cases}
               \nu^{-0.61\pm 0.01} & \text{if}~\lambda\geq 912~\text{\AA},\\
               \nu^{-1.70\pm 0.61} & \text{if}~600~\text{\AA}<\lambda<912~\text{\AA}.\\                
               \end{cases}
\label{eqn:sed}
\end{equation}
Here $f_\nu$ is the flux density, with units of ergs s$^{-1}$
cm$^{-2}$ Hz$^{-1}$, although it is often simply called flux.  Note
that although $f_\nu$ is often the quantity of interest, e.g., when
calculating the hydrogen ionizing emissivity of quasars, what is shown
in spectra is often $f_\lambda$, the flux density
in \emph{wavelength}, which has units of ergs s$^{-1}$
cm$^{-2}$ \AA$^{-1}$.  If the flux density in frequency is power law,
then the flux density in wavelength is also power law.  The power law
index of the flux density in wavelength, $\alpha_\lambda$, is related
to the power law index of the flux density in frequency, $\alpha_\nu$,
by $\alpha_\lambda=-(\alpha_\nu+2)$.  From Equation~(\ref{eqn:sed}),
we can now see that $\alpha_\lambda=-0.3$ in the extreme UV (EUV;
600~{\AA}$<\lambda<$912~{\AA}) and $\alpha_\lambda=-1.39$ in the far
UV (FUV; $\lambda>$912~{\AA}).  The quasar SED that we have adopted is
steep in FUV and shallow in EUV.

How sensible is it to assume that the quasar SED of
Equation~(\ref{eqn:sed}) is universal?  Of the four published
composite quasar SEDs (see references in
\citealt{2015MNRAS.449.4204L}), three agree with
Equation~(\ref{eqn:sed}).  There is one published composite quasar SED
that does not agree with Equation~(\ref{eqn:sed}).  This is the
composite of \citet{2004ApJ...615..135S}, which uses more than 100
quasars at $z<0.1$.  The EUV slope of this SED is $-0.56$.  (This
means that the power law index of the flux density in wavelength is
$-1.44$, so that the spectrum actually becomes even steeper in the
EUV.  See Figure 8 of \citealt{2015MNRAS.449.4204L}.)  The quasars
considered by \citet{2004ApJ...615..135S} are much fainter that those
considered by \citet{2015MNRAS.449.4204L}, which means that it is
possible that faint quasars have a steep EUV spectrum.  So we could
say that for faint quasars,
\begin{equation}
f_\nu\propto\begin{cases}
               \nu^{-0.61\pm 0.01} & \text{if}~\lambda\geq 912~\text{\AA},\\
               \nu^{-0.56\pm 0.61} & \text{if}~600~\text{\AA}<\lambda<912~\text{\AA}.\\                
               \end{cases}
\label{eqn:sed_faint}
\end{equation}
Here, ``faint'' is defined as $M_i(z=2)>-26$.  $M_i(z=2)$ can be
related to the bolometric luminosity of the quasar (see Figure 9 of
\citealt{2015MNRAS.449.4204L}.)

The rate of emission of photons of frequency $\nu$ by quasars can be
written as
\begin{equation}
\dot n_\nu = \int dM_\nu \phi(M_\nu) \frac{L_\nu(M_\nu)}{h\nu}.
\end{equation}
The units of $\dot n_\nu$ are s$^{-1}$~cMpc$^{-3}$~Hz$^{-1}$.  Here
the integral is over some range of magnitudes, the monochromatic
luminosity $L_\nu(M)$ is related to the (absolute AB) magnitude
by \citep{1983ApJ...266..713O}
\begin{equation}
M_\nu = -2.5\log_{10}L_\nu+51.60,
\end{equation}
and $\phi$ is the luminosity function from Equation~(\ref{eqn:dpl}).
The luminosity at 1~Ry is related to that at 1450~\AA\ by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint})
\begin{equation}
  L_{912}=L_{1450}\left(\frac{\nu_{912}}{\nu_{1450}}\right)^{-0.61}=L_{1450}\left(\frac{912}{1450}\right)^{0.61}
\end{equation}
The corresponding volume emissivity is conventionally written
as 
\begin{equation}
\epsilon_\nu = \dot n_\nu h\nu (1+z)^3.
\label{eqn:epsilon}
\end{equation}
This is a \emph{physical} density.  The units of $\epsilon_\nu$ are
erg~s$^{-1}$~pMpc$^{-3}$~Hz$^{-1}$.  This emissivity can now be used
to calculate the ionizing flux and photoionization rate contributed by
quasars.  The frequency dependence above 1~Ry is given by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint}), so that for $\nu >
\nu_{912}$
\begin{equation}
  \epsilon_\nu = \epsilon_{912}\left(\frac{\nu}{\nu_{912}}\right)^\alpha,
  \label{eqn:epsilon_freq}
\end{equation}
where $\alpha=-1.70$ (Equation~\ref{eqn:sed}) or $-0.56$ (Equation~\ref{eqn:sed_faint}).

@@

% Check spellings
% Check notes for more names with add
% Something to add for Joe and Gabor?
We acknowledge useful discussions with James Aird, Manda Banerji,
Tirthankar Roy Choudhury, Xiaohui Fan, Andrea Ferrara, Martin
Haehnelt, Sergey Koposov, Richard McMahon, Ewald Puchwein, Gordon
Richards, and Alberto Rorai for helpful discussions.  GK is supported
by ERC Advanced Grant 320596 `The Emergence of Structure During the
Epoch of Reionization'.

>>>>

\begin{keywords}
dark ages, reionization, first stars -- intergalactic medium --
quasars: general -- galaxies: active
\end{keywords}

%% * Possible outline of paper:
%% 
%% Introduction 
%% Data
%% Method
%%   Individual fits
%%   Global fits
%% Contribution to reionizatin 
%% Conclusions
%% 
%% * Optional possibilities:
%%
%% Low-z photon under-production
%% Bolometric LF
%% Public code description 
%% More global models: PLE, PDE, Trenti, White-Conroy
%% Goodness of fit 
%% BH growth 
%% Comparison with galaxy LF
%% Comparison with previous work

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.  For comparison,
    various other published determinations of the 1~Ry emissivity of
    quasars are also shown in both panels of Figure~\ref{fig:e912}.
    The blue curve in these panels shows the AGN 1~Ry emissivity model
    of \citet{2012ApJ...746..125H}, which is very similar to the model
    of \citet{2007ApJ...654..731H}.  The red symbols with error bars
    show the determination from the fits of
    \citet{2015AA...578A..83G}.  The maroon curve is the published fit
    to the 1~Ry emissivity evolution in the fits of
    \citet{2017MNRAS.466.1160M}.  The green curve is the model of
    \citet{2015ApJ...813L...8M}.}
  \label{fig:e912}
\end{figure*}

@@

The faint-end slope $\beta$ seems to require a break in its evolution
at $z\sim 3$, as discussed in the previous section.  As a result, it
is difficult to model the evolution of $\beta$ using the Chebyshev
expansion of Equation~(\ref{eqn:cbs}).  Instead, we consider a double
power law model for $\beta$ and write
\begin{equation}
  f_3(1+z)=h+\frac{f_0}{10^{a\zeta}+10^{b\zeta}},
\end{equation}
where
\begin{equation}
  \zeta = \log_{10}\left(\frac{1+z}{1+z_0}\right),
\end{equation}
thus resulting in a five-parameter model with parameters $h$, $f_0$,
$z_0$, $a$ and $b$.  The parameters $a$ and $b$ thus determine the low
and high redshift slopes of this evolution, with a break at redshift
$z_0$.  This is similar to the model of \citet{2007ApJ...654..731H},
who also favoured a broken power law model for the evolution of the
faint-end slope of the bolometric luminosity function of quasars.

@@

% Add a table for the best fit parameter values?

@@

We can now calculate the contribution of quasars to the hydrogen
photoionization rate in the intergalactic medium (IGM).  We assume
that all quasars have a universal UV SED, parameterised by
\citet{2015MNRAS.449.4204L} as a power law with a break at 912~{\AA},
\begin{equation}
f_\nu\propto\begin{cases}
               \nu^{-0.61\pm 0.01} & \text{if}~\lambda\geq 912~\text{\AA},\\
               \nu^{-1.70\pm 0.61} & \text{if}~600~\text{\AA}<\lambda<912~\text{\AA}.\\                
               \end{cases}
\label{eqn:sed}
\end{equation}
This fit was derived by \citet{2015MNRAS.449.4204L} from a composite
spectrum of 53 quasars at $z\sim 2.4$.

While four published composite quasar UV SEDs
\citep{2002ApJ...565..773T, 2001AJ....122..549V, 2012ApJ...752..162S,
  2014ApJ...794...75S} agree with Equation~(\ref{eqn:sed}), the
composite of \citet{2004ApJ...615..135S}, which uses more than 100
quasars at $z<0.1$, is somewhat steeper, with a EUV slope of $-0.56$.
The quasars considered by \citet{2004ApJ...615..135S} are fainter
($M_i(z=2)>-26$) than those considered by \citet{2015MNRAS.449.4204L},
which suggest that faint quasars may have a steeper EUV spectrum.  We
therefore also consider an SED of the form
\begin{equation}
f_\nu\propto\begin{cases}
               \nu^{-0.61\pm 0.01} & \text{if}~\lambda\geq 912~\text{\AA},\\
               \nu^{-0.56\pm 0.61} & \text{if}~600~\text{\AA}<\lambda<912~\text{\AA}.\\                
               \end{cases}
\label{eqn:sed_faint}
\end{equation}
for qsos with $M_{1450}<-23$.

The rate of emission of photons of frequency $\nu$ by quasars can be
written as
\begin{equation}
\dot n_\nu = \int dM_\nu \phi(M_\nu) \frac{L_\nu(M_\nu)}{h\nu}.
\end{equation}
Here the integral is over a suitable range of magnitudes, the
monochromatic luminosity $L_\nu(M)$ is related to the (absolute AB)
magnitude by \citep{1983ApJ...266..713O}
\begin{equation}
M_\nu = -2.5\log_{10}L_\nu+51.60,
\end{equation}
and $\phi$ is the luminosity function from Equation~(\ref{eqn:dpl}).
The luminosity at 1~Ry is related to that at 1450~\AA\ by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint})
\begin{equation}
  L_{912}=L_{1450}\left(\frac{\nu_{912}}{\nu_{1450}}\right)^{-0.61}=L_{1450}\left(\frac{912}{1450}\right)^{0.61},
\end{equation}
depending on the qso luminosity.  The corresponding physical volume
emissivity is 
\begin{equation}
\epsilon_\nu = \dot n_\nu h\nu (1+z)^3.
\label{eqn:epsilon}
\end{equation}
Figure~\ref{fig:e912} shows the resultant 1~Ry emissivity when the
luminosity function is integrated down to $M_{1450}=-20$ (right panel)
and when it is integrated down to $M_{1450}=-18$ (left panel).
Similar to the previous figures, the black open circles in this figure
show the result form individual redshift bins, while the black curve
shown the result from best-fit global model.  Yellow curves show a
random sample from the global posterior distribution.  For comparison,
various other published determinations of the 1~Ry emissivity of
quasars are also shown in both panels of Figure~\ref{fig:e912}.  The
blue curve in these panels shows the AGN 1~Ry emissivity model of
\citet{2012ApJ...746..125H}, which is very similar to the model of
\citet{2007ApJ...654..731H}.  The red symbols with error bars show the
determination from the fits of \citet{2015AA...578A..83G}.  The maroon
curve is the published fit to the 1~Ry emissivity evolution in the
fits of \citet{2017MNRAS.466.1160M}.  The green curve is the model of
\citet{2015ApJ...813L...8M}.  Note that the integration limits for
these other models are not homogeneous.  \citet{2017MNRAS.466.1160M}
integrate the luminosity function down to $M_{1450}=-19$, while
\citet{2015AA...578A..83G} integrate down to $M_{1450}=-18$.
\textbf{[What are the limits for HM12 and MH15?]}

The 1~Ry emissivity increases rapidly from $z=0$ up to $z=2$ and then
stays roughly constant until up to $z=3.5$.  At higher redshifts, due
to the relatively steeper faint end slope of the luminosity function,
the integration limit makes a large difference in the resultant
emissivity.  For the conservative case, in which we integrate under
the luminosity function down to $M_{1450}=-20$, the emissivity drops
rapidly at $z>3.5$, and is only about 10--20\% higher than that in the
\citet{2012ApJ...746..125H} model at $z\sim 6$.  However, in the model
in which we integrate down to $M_{1450}=-18$, the emissivity is higher
by a factor of almost 10 than that in the \citet{2012ApJ...746..125H}
model at $z\sim 6$.  In this case, our derived emissivities are
consistent with the values published by \citet{2015AA...578A..83G} at
$z=4$--$6$, although they still fall short by a factor of $\sim 2$.
Our derived emissivities are significantly smaller at $z>3.5$ than the
model of \citet{2015ApJ...813L...8M}.

The 1~Ry emissivity can now be used to calculate the ionizing flux and
photoionization rate contributed by quasars.  The frequency dependence
above 1~Ry is given by Equation~(\ref{eqn:sed}) or
(\ref{eqn:sed_faint}), so that for $\nu > \nu_{912}$
\begin{equation}
  \epsilon_\nu = \epsilon_{912}\left(\frac{\nu}{\nu_{912}}\right)^\alpha,
  \label{eqn:epsilon_freq}
\end{equation}
where $\alpha=-1.70$ (Equation~\ref{eqn:sed}) or $-0.56$ (Equation~\ref{eqn:sed_faint}).

@@

% Check spellings
% Check notes for more names with add
% Something to add for Joe and Gabor?
We acknowledge useful discussions with James Aird, Manda Banerji,
Tirthankar Roy Choudhury, Xiaohui Fan, Andrea Ferrara, Martin
Haehnelt, Sergey Koposov, Richard McMahon, Ewald Puchwein, Gordon
Richards, and Alberto Rorai.  GK is supported by ERC Advanced Grant
320596 `The Emergence of Structure During the Epoch of Reionization'.
